\mode*

% What's the problem?
% Why is it a problem? Research gap left by other approaches?
% Why is it important? Why care?
% What's the approach? How to solve the problem?
% What's the findings? How was it evaluated, what are the results, limitations, 
% what remains to be done?

% XXX Summary
\emph{Summary:}
In this learning session we will cover the foundations of security.
By this we mean what security is all about, \eg what types of properties we are 
interested in and what we want to achieve in our security work.
We will also introduce the scientific method and, particularly, how this can be 
applied in the area of security to create new knowledge.

Finally, we will cover usability and how users' weaknesses affects security.
There are many ways to attack systems through their human operators.
We must consider usability when designing secure systems.

% XXX Motivation and intended learning outcomes
\emph{Intended learning outcomes:}
After this session you should be able:
\begin{itemize}
  \item to \emph{understand} the what security is generally about.
  \item to \emph{differentiate} which types of scientific methods are 
    appropriate to answer a given question.
  \item to \emph{adopt} an adversarial thinking for situtions involving humans.
  \item to \emph{incorporate} basic psychology in the design of a system to 
    increase its security.
\end{itemize}

% XXX Prerequisites
%\emph{Prerequisites:}
%\dots

% XXX Reading material
\emph{Reading:}
You should read Gollmann's chapter on \enquote{Foundations of Computer 
  Security}~\cite[Chap.\ 3]{Gollmann2011cs}.
There he attempts at a definition of Computer Security and related terms, \eg 
confidentiality, integrity, and availability, which we need for our treatment of 
the topic.
Anderson also covers this in Chapter 1 of~\cite{Anderson2008sea}.
He also treats a wider area than just \emph{computer} security, which is good 
for us, he covers many aspects of security in different examples.

For the introduction to the scientific method you should read 
\citetitle{HowToDesignSecurityExperiments}~\cite{HowToDesignSecurityExperiments}.
This paper discusses the scientific method of (parts of) the security field.
For a more in-depth reflection on the state of security as a scientific pursuit, 
we recommend
\citetitle{SecurityAsAScience}~\cite{SecurityAsAScience}.

\Citeauthor{Anderson2008sea} gives a short summary of the psychology of users, 
their strengths and weaknesses, and how usability affects security in Chapter 2 
\enquote{Usability and Psychology} of 
\citetitle{Anderson2008sea}~\cite{Anderson2008sea}.
