% $Id$
% Author:  Daniel Bosk <daniel.bosk@miun.se>
\documentclass[svv,addpoints]{miunexam}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage[swedish,english]{babel}
\usepackage[hyphens]{url}
\usepackage{hyperref}
\usepackage{color}
\usepackage{prettyref,varioref}
\usepackage{subfigure}
\usepackage{amsmath,amssymb}
\usepackage{listings}
\usepackage{authblk}

\usepackage{csquotes}
\MakeBlockQuote{<}{|}{>}
\EnableQuotes

\usepackage[natbib,style=alphabetic,maxbibnames=99,backend=bibtexu]{biblatex}
\addbibresource{literature.bib}

%\printanswers

\examtype{Final exam}
\courseid{DT145G}
\course{Computer Security}
\date{2016-06-03}
\author{%
  Daniel Bosk
}
\affil{%
  Department of Information and Communication Systems,\\
  Mid Sweden University, SE-851\,70 Sundsvall\\
  Email: \href{mailto:daniel.bosk@miun.se}{daniel.bosk@miun.se}\\
  Phone: 010-142\,8709
}

\DeclareMathOperator{\hmac}{HMAC}
\DeclareMathOperator{\xor}{\oplus}
\DeclareMathOperator{\concat}{||}

\begin{document}
\maketitle
\thispagestyle{foot}

\section*{Instructions}
\label{sec:Instructions}
Carefully read the questions before you start answering them.
Note the time limit of the exam and plan your answers accordingly.
Only answer the question, do not write about subjects remotely related to the
question.

Write your answers on separate sheets, not on the exam paper.
Only write on one side of the sheets.
Start each question on a new sheet.
Do not forget to \emph{motivate your answers.}

Make sure you write your answers clearly, if I cannot read an answer the answer
will be awarded no points---even if the answer is correct.
The questions are \emph{not} sorted by difficulty.

\begin{description}
  \item[Time] 5 hours.
  \item[Aids] Dictionary.
  \item[Maximum points] \numpoints
  \item[Questions] \numquestions
\end{description}

\subsection*{Preliminary grades}
The following grading criteria applies:
E \(\geq 50\%\),
D \(\geq 60\%\),
C \(\geq 70\%\), 
B \(\geq 80\%\),
A \(\geq 90\%\).


\clearpage
\section*{Questions}
The questions are given below.
They are not given in any particular order.

\begin{questions}
\question\label{q:usability:E:C:A}
  Human psychology is important in security.
  It is used in both security usability and social engineering.
  \begin{parts}
    \part[2] Give an overview of why psychology is important in 
    security.

    \begin{solution}
      Då systemen vi är beroende av och som ska upprätthålla vår säkerhet 
      handhas av människor blir psykologin genast viktig.
      Vi behöver psykologin inom säkerhetsområdet för att kunna ta hänsyn till 
      hur människor fungerar när vi konstruerar säkerhetssystem.
      Exempelvis, om vi gör ett system för komplext och användaren tycker att 
      komplexiteten är onödig, då kommer denne användare att aktivt försöka att 
      ta sig runt systemet --- kanske genom att skriva upp långa lösenord 
      istället för att lära sig dem utantill.

      Om vi däremot tar hänsyn till användarnas kognitiva begränsningar, då kan 
      vi konstruera system som både är säkra och enkla att använda.
    \end{solution}

    \part[4] Give an example of an attack which exploits weaknesses in human 
    psychology.
    Also explain why it works.

    \begin{solution}
      En psykologibaserad attack utnyttjar svagheter hos användarna för att ta 
      sig runt ett säkerhetssystem, det är alltså inte säkerhetssystemen som 
      angrips.

      Ett exempel på en sådan attack kan vara att en användare får ett e-brev 
      som till synes är från banken och som innehåller en länk till en 
      inloggningssida, kallat nätfiske.
      Brevet kan be användaren att uppdatera någonting hos banken via internet.
      Ett förfarande beskrivs och sedan läggs till \enquote{eller klicka på 
        länken}.
      Med en förfarande som låter som att det kan ta fem till tio klick kommer 
      användaren sannolikt att välja enklicksalternativet.
      Notera att förfarandet måste vara korrekt för banken medan länken är till 
      en phishingsida.
      Utformandet kan leda till vad litteraturen~\cite[s. 23]{Anderson2008sea} 
      kallar \emph{\foreignlanguage{english}{capture errors}}, att användaren 
      använder ett invant beteende: i detta fall att användaren klickar på 
      direktlänkar.

      Därutöver försöker nätfiskaren att få användaren att tillämpa fel regler 
      i situationen.
      Exempelvis, användaren kanske (omedvetet) lägger större vikt vid att ett 
      hänglås syns i webbläsaren för säker anslutning än att bankens namn är 
      rätt stavat i URL:en.
      Även att bankens namn finns med någonstans i URL:en kan vara en 
      tillräckligt stark regel för att användaren ska undvika att detektera den 
      felaktiga fiske-URL:en.
    \end{solution}
  \end{parts}


  
\question\label{q:accountability:E:C}
  Separation of duties is a core concept for security.
  \begin{parts}
    \part[2] Describe the two types of separation of duties.
    \part[1] What is the main reason for separation of duties?
  \end{parts}

  \begin{solution}
    There are two types of separation of duties:
    dual control and functional separation.
    Dual control means that two or more subjects must act together (at the same 
    time) to authorize a transaction.
    Functional separation means that several functions are needed to authorize 
    a transaction---e.g.~create a transaction and verify it---and one subject 
    is not allowed to do both functions.

    The reason for separation of duties to make it impossible for one malicious 
    subject to compromise a system.
    With separation of duties the malicious subject must persuade one or more 
    other subjects to collude.
  \end{solution}


  
\question[4]\label{q:trustcomp}
% examgen: trustcomp:E:C:A
Give an example of a DRM system, the idea behind it and why it works or not.

\begin{solution}
  Hardware dongles:
  You have a hardware dongle attached to the computer, the software can then 
  communicate with the dongle.
  The idea is that the software can be copied easily, but the dongle cannot.
  Thus the software can only run in as many instances as there are hardware 
  dongles.

  The hardware dongle can be simulated by other software in many cases.
  For the software to be able to tell the dongle and the simulated dongle 
  apart, it must be able to trust the operating system --- thus it needs to 
  verify the integrity of the operating system, which in turn requires special 
  hardware.
  The alternative approach is that the dongle is more sophisticated, e.g.\ that 
  it uses unforgeable digital signatures as output.
  In this case we instead modify the software itself, so that it simply skips 
  the checks with the dongle (e.g.\ the signature verification always returns 
  true).
\end{solution}



\question\label{q:foundations:E}
  Define the following terms:
  \begin{parts}
    \part[1] Trusted
    \part[1] Trustworthy
    \part[1] Secrecy
    \part[1] Confidentiality
    \part[1] Privacy
    \part[1] Integrity
    \part[1] Authenticity
  \end{parts}

  \begin{solution}
    \citet[ss.\ 13--14]{Anderson2008sea} definierar begreppen enligt följande:
    \begin{description}
      \item[Pålitlighet] Ett system eller principal som innehar pålitlighet 
        (\foreignlanguage{english}{is trusted}) är ett system eller principal 
        som kan bryta din säkerhetspolicy.

      \item[Pålitlig] Ett system eller principal som är pålitlig 
        (\foreignlanguage{english}{is trustworthy}) är ett system eller 
        principal som inte kommer att misslyckas.
        (Den kommer alltså inte att bryta din säkerhetspolicy.)

        Ett exempel för att illustrera skillnaden ges av följande citat: 
        \blockcquote[s.\ 13]{Anderson2008sea}{%
          if an NSA employee is observed in a toilet stall at Baltimore 
          Washington airport selling key material to a Chinese diplomat, then 
          (assuming his operation was not authorized) we can describe him as 
          \enquote{trusted but not trustworthy}%
        }.

      \item[Sekretess] Sekretess är en teknisk term för effekten av en mekanism 
        som begränsar antalet principals som kan ta del av information.

      \item[Konfidentialitet] Konfidentialitet syftar till att tillhandahålla 
        sekretess för andra principals hemliga information.

      \item[Personlig integritet] Detta är förmågan eller rätten att kunna 
        skydda sin personliga information.
        Det gäller alltså bara individer, exempelvis företag har ingen 
        personlig integritet.

      \item[Integritet] Detta är en teknisk term för egenskapen att data 
        förblir oförändrat, eller, om förändring sker ska den inte förbli 
        obemärkt.

      \item[Autenticitet] Detta begrepp innefattar integritet och fräshhet.
        Om kommunikation spelas in och sedan spelas upp vid ett annat 
        tillfälle, då kommer integriteten att ha bevarats men inte fräshheten 
        --- alltså är en återuppspelning inte autentisk.
    \end{description}
    Dessa definitioner stämmer även överens med RFC 4949~\cite{rfc4949}.
  \end{solution}


  
\question[4]\label{q:sidechannels}
% examgen: sidechannels:E
Give an example of a side-channel attack and motivate why it is a side channel.

\begin{solution}
  A side channel is an unintended channel emitting information which is due 
  to physical implementation flaws and not theoretical weaknesses or forcing 
  attempts.

  (2 points) Extracting the secret key from a device by measuring energy 
  consumption or electromagnetic emissions while the device performs 
  computations using the secret key.

  (1 point) This is a side channel since it relies on a weakness in the 
  hardware implementation.
  (1 point) It is further an active attack since we might need the device to 
  perform operations on certain ciphertexts (or plaintexts).
\end{solution}


%
\question\label{q:crypto:foundations:E}
  Explain the following terms:
  \begin{parts}
    \part[1] Confidentiality
    \part[1] Integrity
    \part[1] Availability
    \part[1] Accountability
    \part[1] Non-Repudiation
  \end{parts}

  \begin{solution}
    See~\cite{Gollmann2011cs} and~\cite{Anderson2008sea} for definitions.
  \end{solution}


  
\question\label{q:passwd:infotheory:E:C:A}
  You are asked to estimate some password policies.
  The policies are the following:
  \begin{description}
    \item[basic12]
      At least 12 characters consisting of upper and lower case, and numbers.
    \item[randswedict4]
      Randomly choose four words from the Dictionary of the Swedish Language 
      (SAOL).
      This dictionary contains approximately 125\,000 words.
  \end{description}
  You should answer the following:
  \begin{parts}
    \part[4] Estimate the entropy for the password policies.
    (You may rely on the results in certain published research papers discussed 
    in the course for certain estimates.)
    \part[2] Decide how suitable they are for use in the home environment.
    \part[2] Decide how suitable they are for use in a web application.
  \end{parts}
  Note that you will not get any points without a motivation.


  
\question[3]\label{q:software}
% examgen: software:E:C:A
Give an example where \enquote{data} can be mistaken for \enquote{code}.

\begin{solution}
  Shell scripting is an easy example.
  Here you can store part of the code in variables, the simply substitute them.
  Consider the following \texttt{/bin/echo -e \$\{1\}}.
  The variable \texttt{\$\{1\}} will be substituted and the result will be 
  interpreted as code.
\end{solution}



\question\label{q:passwd:auth:E}
  Explain the following terms:
  \begin{parts}
    \part[1] Brute force
    \part[1] Dictionary attack
    \part[1] Hash table
    \part[1] Social engineering
    \part[1] Two-factor authentication
    \part[1] Phishing
  \end{parts}


%  \%question\label{q:msb}
%  Explain the following terms:
%  \begin{parts}
%    \part[1] GAP-analys
%    \part[1] LIS
%    \part[1] säkerhetspolicy
%    \part[1] nulägesanalys
%    \part[1] ISO 27000
%  \end{parts}


  
\end{questions}


\printbibliography
\end{document}

